### 2017-04-30 fhb-meta.Rmd
corrections

---
title: "Meta-analysis of fungicide efficacy for FHB control in Brazil"
output: github_document
---

```{r, warning=FALSE, include=FALSE}

# Load packages 
library(tidyverse)
library(metafor)
library(ggthemes)
library(cowplot)
library(scales)
library(knitr)
library(nloptr)
library(minqa)
```

## Introduction

## Import data

```{r}

# import data

# FHB index data
fhb_sev <- read.csv("fhb_sev.csv", sep = ",", h = T)

# Yield data
fhb_yield <- read.csv("fhb_yield.csv", sep = ",", h = T)
```

## Prepare variables for meta-analysis

### Effect size

```{r}

# log of FHB index
fhb_sev$yi <- log(fhb_sev$sev)

# log of yield
fhb_yield$yi <- log(fhb_yield$yield)

# difference in yield between treatment and non-treated check
fhb_yield$D <-fhb_yield$yield - fhb_yield$yield_check

# yield and 
fhb_yield$yi2 <- fhb_yield$yield

```

### Sampling variance

```{r}

# Sampling variance for the log of FHB index
fhb_sev$vi <- with(fhb_sev, V_sev / (n * sev^2))

# Sampling variance for the log of yield
fhb_yield$vi <- with(fhb_yield, V_yield / (n * yield^2))

# Sampling variance for yield
fhb_yield$vi2 <- fhb_yield$V_yield/fhb_yield$n # multivariate approach

```

### Create treatment variables

```{r}
# Check the number of entries by fungicide (AI: active ingredient) and number of sprays
table(fhb_sev$AI, fhb_sev$n_spray2)

# Create treatment variable based on fungicide name and number of sprays 
fhb_sev <- fhb_sev %>% 
  mutate(AI_nspray2 = paste(AI, n_spray2,sep='_')) %>% 
  filter(AI_nspray2 != "CARB_1") %>% 
  filter(trial, length(trial)>1)

fhb_yield <- fhb_yield %>% 
  mutate(AI_nspray2 = paste(AI, n_spray2,sep='_')) %>% 
  filter(AI_nspray2 != "CARB_1") %>% 
  filter(trial, length(trial)>1)

```

## ID variable for each entry within a trial

```{r}
fhb_sev$id <- 1:nrow(fhb_sev)
fhb_yield$id <- 1:nrow(fhb_yield)
```

## Descriptive analysis for yield

```{r}
# Number of entries by fungicide and number of sprays
kable(table(fhb_yield$AI, fhb_yield$n_spray2), format="pandoc")

# Number of unique trials
length(unique(fhb_yield$trial))

# Number of unique studies
length(unique(fhb_yield$study))

# Number of trial by year
fhb_trial <- fhb_yield %>% 
  group_by(trial) %>% 
  filter(row_number() ==1)

kable(data.frame(table(fhb_trial$year)), format="pandoc")

nrow(table(fhb_trial$year))

# Number of trial by location
kable(data.frame(table(fhb_trial$location)), format="pandoc")

nrow(table(fhb_trial$location))#Ponta Grossa repete (??)

# Number of trial by publication
kable(data.frame(table(fhb_trial$publication)), format="pandoc")

# Number of trial by state
kable(data.frame(table(fhb_trial$state)), format="pandoc")

# entries with D value lower than zero
negat_D <- fhb_yield %>% 
  filter(D < 0)

```

## Calculate coefficient of variation of FHB index and yield

```{r}
# CV for FHB index
## All entries
cv <- (sd(fhb_sev$sev)/mean(fhb_sev$sev))*100
cv

## Only non-treated check
cv <- (sd(fhb_sev$sev_check)/mean(fhb_sev$sev_check))*100
cv

### CV for yield

## All entries
cv <- (sd(fhb_yield$yield)/mean(fhb_yield$yield))*100
cv

## For control
cv <- (sd(fhb_yield$yield_check)/mean(fhb_yield$yield_check))*100
cv
```

## Network meta-analysis

### Mean log of FHB index

```{r, message=FALSE, warning=FALSE}
fhb_mv_AI <- rma.mv(yi, vi, 
                    mods = ~ AI_nspray2, 
                    method="ML",
                    random = list(~ AI_nspray2 | trial, ~1 | id), 
                    struct="UN", data=fhb_sev)

# Linear contrasts between pairs of treatments
anova(fhb_mv_AI, L=rbind(c(0,1,-1,0,0), 
                         c(0,0,-1,1,0),
                         c(0,0,-1,0,1),
                         c(0,0,0,-1,1)))

```

## Control efficacy

Back-transform (exponential) the difference of the logs to obtain the response ratio and further calculate the percent reduction in FHB index relative to the check treatment (control efficacy) for each fungicide.

```{r}
results_AI <- data.frame(cbind((1-exp(fhb_mv_AI$b))*100, 
                              (1-exp(fhb_mv_AI$ci.lb))*100,
                              (1-exp(fhb_mv_AI$ci.ub))*100))
results_AI
```

## Inconsistency test

```{r, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}

### Number of trials per design (cons_group)
table(fhb_sev$cons_group)

### Model testing interaction treat x design
fhb_mv_AI_cons <- rma.mv(yi, vi, 
                         mods = ~ AI_nspray2 * cons_group, 
                         method="ML",
                         random = ~ AI_nspray2 | trial/cons_group, 
                         struct="UN", 
                         data=fhb_sev)

summary(fhb_mv_AI_cons)

results_I<- data.frame(cbind((exp(fhb_mv_AI$b)-1)*100, 
                             (exp(fhb_mv_AI$ci.lb)-1)*100,
                             (exp(fhb_mv_AI$ci.ub)-1)*100))

results_I
anova(fhb_mv_AI_cons, btt=11:20) #p-val = 0.9999

```

## Moderator variables

### FHB index base

```{r, message=FALSE, warning=FALSE}
# create the binary variable (dis_press)
fhb_sev$dis_press <- ifelse(as.numeric(fhb_sev$sev_check) >7.0, 2, 1)

# summarize number of trials per category of dis_press
table(fhb_sev$AI_nspray2, fhb_sev$dis_press)

# Test effect of moderator 
fhb_mv_AI_sev_check <- rma.mv(yi, vi, 
                              mods = ~ AI_nspray2*factor(dis_press),
                              method="ML",
                              random = list(~ AI_nspray2 | trial, ~1 | id),
                              struct="UN", 
                              data=fhb_sev)

# Contrast levels of moderators
anova(fhb_mv_AI_sev_check,btt=7:10)

```

### Yield base

```{r, message=FALSE, warning=FALSE}
# create binary variable (yield_class)
summary(fhb_sev$yield_check) # Median = 2993; Mean = 2915
fhb_sev$yield_class <- ifelse(fhb_sev$yield_check > 3000, 2, 1)

# check number of trials per yield_class category
table(fhb_sev$yield_class, fhb_sev$AI_nspray2)

# test the effect of moderator
fhb_mv_AI_yield_class <- rma.mv(yi, vi, 
                                mods = ~ AI_nspray2*factor(yield_class),
                                method="ML",
                                random = list(~ AI_nspray2 | trial, ~1 | id), 
                                struct="UN", 
                                data=fhb_sev)

anova(fhb_mv_AI_yield_class, btt=7:10)

```

### Year as continuous

```{r, message=FALSE, warning=FALSE}
# Moderator year (continuous)
# Number of entries by fungicide and year
table(fhb_sev$AI_nspray2, fhb_sev$year)

fhb_mv_AI_year <- rma.mv(yi, vi, 
                         mods = ~ AI_nspray2*as.numeric(year), 
                         method="ML",
                         random =list(~ AI_nspray2 | trial, ~1 | id),
                         struct="UN", 
                         data=fhb_sev)

anova(fhb_mv_AI_year, btt=7:10)

```

### Overall percent yield increase by treament

```{r}
### AI_nspray2 (AI and number of sprays)

fhb_mv_AI <- rma.mv(yi, vi, 
                    mods = ~ AI_nspray2, 
                    method="ML",random = list(~ AI_nspray2 | trial, ~1 | id), 
                    struct="UN", data=fhb_yield)

summary(fhb_mv_AI)

results_I<- data.frame(cbind((exp(fhb_mv_AI$b)-1)*100, 
                             (exp(fhb_mv_AI$ci.lb)-1)*100,
                             (exp(fhb_mv_AI$ci.ub)-1)*100))

results_I

# Linear contrasts between treatments

anova(fhb_mv_AI, L=rbind(c(0,1,-1,0,0), 
                         c(0,0,1,-1,0),
                         c(0,0,1,0,-1),
                         c(0,0,0,-1,1)))

```

### Overall absolute yield increase

```{r}
### by fungicide treatment

fhb_mv_AI_D <- rma.mv(yi2, vi2,
                      mods = ~ AI_nspray2, 
                      method="ML",random = list(~ AI_nspray2 | trial, ~1 | id),                       struct="UN", 
                      data=fhb_yield, 
                      control = list(optimizer="nlm"))
summary(fhb_mv_AI_D)

# Linear contrasts between treatments

anova(fhb_mv_AI_D, L=rbind(c(0,1,-1,0,0), 
                         c(0,0,1,-1,0),
                         c(0,0,1,0,-1),
                         c(0,0,0,-1,1)))
```

### Inconsistency test for yield data
```{r}

### Groups were inclued directly in the spreadsheet
table(fhb_yield$cons_group)

### Model 
fhb_mv_AI_cons <- rma.mv(yi2, vi2, 
                         mods = ~ AI_nspray2 * cons_group, 
                         method="ML",
                         random = ~ AI_nspray2 | trial/cons_group, 
                         struct="UN", data=fhb_yield)

summary(fhb_mv_AI_cons)

results_I<- data.frame(cbind((exp(fhb_mv_AI$b)-1)*100, 
                             (exp(fhb_mv_AI$ci.lb)-1)*100,
                             (exp(fhb_mv_AI$ci.ub)-1)*100))
results_I

#anova(fhb_mv_AI_cons, btt=14:26)

```

## Visualize the data

###Boxplot of severity by treatment

```{r}
ggplot(fhb_sev, aes(AI_nspray2, sev))+
  geom_boxplot(size = 0.5, outlier.shape = NA)+ 
  geom_jitter(width=0.1, shape=1, size=2.5, color="gray50")+
  scale_x_discrete(labels=c("CHECK","CARB 2x", "PROP 2x" ,"TEBU 1x","TEBU 2x")) +
  theme_classic()+
  xlab("")+
  ylab("FHB index (%)")+ 
  scale_y_continuous(breaks=c(0, 10,  20,  30, 40, 50, 60, 70, 80, 90, 100), limits=c(0,100))

```

### Boxplot yield

```{r}

ggplot(fhb_yield, aes(AI_nspray2, yield))+
  geom_boxplot(size = 0.5, outlier.shape = NA)+ geom_jitter(width=0.1, shape=1, size=2.5, color="gray50")+
  scale_x_discrete(labels=c("CHECK","CARB 2x", "PROP 2x" ,"TEBU 1x","TEBU 2x")) +
  theme_cowplot()+
  xlab("")+
  ylab("Yield (kg/ha)")+
  scale_y_continuous(breaks=c(0, 1000,  2000, 3000, 4000, 5000, 6000))

```

### Plot D

```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(plyr)
fhb_yield$se <- sqrt(fhb_yield$vi2/sqrt(fhb_yield$n))

sum<-ddply(.data=fhb_yield, 
               .(AI_nspray2), 
               summarize, 
               n=paste("n =", length(AI_nspray2)))
sum1<-sum %>% filter(AI_nspray2 != "AACHECK_0", AI_nspray2!="PROP_2",AI_nspray2 !="TEBU_1",AI_nspray2 !="TEBU_2")

##### CARB ########

plot_D_carb <- fhb_yield %>% group_by(trial) %>% filter(AI_nspray2 != "AACHECK_0", AI_nspray2!="PROP_2",AI_nspray2 !="TEBU_1",AI_nspray2 !="TEBU_2") %>% 
  ggplot(., aes(reorder(treatment, D), D)) +
  geom_bar(stat="identity")+
  geom_errorbar(width=0.5, size= 0.3, aes(ymax = D + se, ymin=D - se))+
  geom_text(data=sum1, aes(x=5, y=1200,label=n), inherit.aes=FALSE, parse=FALSE)+
  geom_text(aes(x=5, y=1400,label="CARB 2x"), inherit.aes=FALSE, parse=FALSE)+
  geom_hline(aes(yintercept=0), size = 0.1)+
  scale_y_continuous(breaks=c(-250,0, 250,  500, 750, 1000, 1250, 1500))+
  ylab("D (kg/ha)")+ xlab("Study treatments (sorted) ") +
  theme(axis.text = element_text(size=11),
        #axis.title.x=element_blank(), 
        #axis.ticks.x=element_blank(),
        axis.text.x=element_blank(),
        legend.position="none")

######PROP######

sum2<-sum %>% filter(AI_nspray2 != "AACHECK_0", AI_nspray2!="CARB_2",AI_nspray2 !="TEBU_1",AI_nspray2 !="TEBU_2")

plot_D_prop <- fhb_yield %>% group_by(trial) %>% filter (AI_nspray2 != "AACHECK_0", AI_nspray2!="CARB_2",AI_nspray2 !="TEBU_1",AI_nspray2 !="TEBU_2") %>% 
  ggplot(., aes(reorder(treatment, D), D)) +
  geom_bar(stat="identity")+
  geom_errorbar(width=0.5, size= 0.3, aes(ymax = D + se, ymin=D - se))+
  geom_text(data=sum2, aes(x=4, y=1200,label=n), inherit.aes=FALSE, parse=FALSE)+
  geom_text(aes(x=4, y=1400,label="PROP 2x"), inherit.aes=FALSE, parse=FALSE)+
  geom_hline(aes(yintercept=0), size = 0.1)+
  scale_y_continuous(breaks=c(-250,0, 250,  500, 750, 1000, 1250, 1500))+
  ylab(" ")+ xlab("Study treatments (sorted) ") +
  theme(axis.text = element_text(size=11),
        axis.text.x=element_blank(),
        legend.position="none")

##### TEBU########

sum3<-sum %>% filter(AI_nspray2 != "AACHECK_0", AI_nspray2!="CARB_2",AI_nspray2 !="PROP_2",AI_nspray2 !="TEBU_2")

plot_D_tebu <- fhb_yield %>% group_by(trial) %>% filter (AI_nspray2 != "AACHECK_0", AI_nspray2!="CARB_2",AI_nspray2 !="PROP_2",AI_nspray2 !="TEBU_2") %>% 
  ggplot(., aes(reorder(treatment, D), D)) +
  geom_bar(stat="identity")+
 geom_errorbar(width=0.5, size= 0.3, aes(ymax = D + se, ymin=D - se))+
  geom_text(data=sum3, aes(x=5, y=1300,label=n), inherit.aes=FALSE, parse=FALSE)+
  geom_text(aes(x=5, y=1500,label="TEBU 1x"), inherit.aes=FALSE, parse=FALSE)+
  geom_hline(aes(yintercept=0), size = 0.1)+
  scale_y_continuous(breaks=c(-250,0, 250,  500, 750, 1000, 1250, 1500))+
  ylab("D (kg/ha)")+ xlab("Study treatments (sorted) ") +
  labs(fill="Number of sprays" )+
    theme(axis.text = element_text(size=11),
        axis.text.x=element_blank(),
        legend.position="top")

sum4<-sum %>% filter(AI_nspray2 != "AACHECK_0", AI_nspray2!="CARB_2",AI_nspray2 !="PROP_2",AI_nspray2 !="TEBU_1")

plot_D_tebu2 <- fhb_yield %>% group_by(trial) %>% filter (AI_nspray2 != "AACHECK_0", AI_nspray2!="CARB_2",AI_nspray2 !="PROP_2",AI_nspray2 !="TEBU_1") %>% 
  ggplot(., aes(reorder(treatment, D), D)) +
  geom_bar(stat="identity")+
  geom_errorbar(width=0.5, size= 0.3, aes(ymax = D + se, ymin=D - se))+
  geom_text(data=sum4, aes(x=5, y=1200,label=n), inherit.aes=FALSE, parse=FALSE)+
  geom_text(aes(x=5, y=1400,label="TEBU 2x"), inherit.aes=FALSE, parse=FALSE)+
  geom_hline(aes(yintercept=0), size = 0.1)+
  scale_y_continuous(breaks=c(-250,0, 250,  500, 750, 1000, 1250, 1500))+
  ylab(" ")+ xlab("Study treatments (sorted) ") +
  labs(fill="Number of sprays" )+
    theme(axis.text = element_text(size=11),
        axis.text.x=element_blank(),
        legend.position="top")

combo1 <- plot_grid(plot_D_carb, plot_D_prop,labels=c('', ''),  ncol = 2)

combo2 <- plot_grid(plot_D_tebu, plot_D_tebu2,labels=c('', ''),  ncol = 2)

plot_grid(combo1, combo2,labels=c('', ''),  ncol = 1)

```