### 2017-07-24 code.Rmd

---
title: ''
output:
  html_document:
    css: my-style.css
    df_print: paged
    theme: flatly
    highlight: default
    toc: true
    toc_float: 
     collapsed: true
     smooth_scroll: false
---

## Introduction

We conducted a two-stage multi-treatment or network meta-analysis for the simultaneous analysis of all treatments of interest and which co-occur in the same trial. This approach allows direct comparisons of treatments with each other, and takes into account all the correlations (see Madden et al. 2016 for more details).

The most common form of effect size used in traditional meta-analysis are based on the contrasts of the treatment of interest with a common reference (e.g. control treatment), such as mean difference, response ratio, etc. This is known as the conditional modeling approach, also named contrast-based meta-analysis. An alternative and simpler approach, which is commonly used in plant pathology, is to fit a two-way linear mixed model directly to the treatment means. This is know as the unconditional modeling approach or arm-based meta-analysis.

## Import data

Let's first import the data and load the packages needed for the analyses.

```{r}
fhb_sev <- read.csv("fhb_sev.csv", sep = ",", h = T)
fhb_yield <- read.csv("fhb_yield.csv", sep = ",", h = T)
```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
# Load packages 
library(tidyverse)
library(metafor)
library(ggthemes)
library(cowplot)
library(scales)
library(knitr)
library(nloptr)
library(minqa)
library(DT)
```

## Effect-sizes

Here, we used the latter approach fitted directly to the log of the means (for both FHB index and yield) for further obtaining the relative effect (control efficacy and yield response). For yield, we fitted the model directly to the mean yield of treatments to further estimate the yield response, or the difference (D) from using the fungicides. The D was calculated for each treatment within a study for plotting purposes, but the meta-analyitic estimate of D was obtained from the difference of the estimates by mixed model fitted directly to the treatment means.

```{r}
# log of FHB index
fhb_sev$yi <- log(fhb_sev$sev)
hist(fhb_sev$yi)

# log of yield
fhb_yield$yi <- log(fhb_yield$yield)
hist(fhb_yield$yi)

# difference (D) in yield between treatment and non-treated check
fhb_yield$D <-fhb_yield$yield - fhb_yield$yield_check
hist(fhb_yield$D)
```

## Sampling variance

Let's create the sampling variance variables `vi` for the respective effect-sizes.

```{r}
# Sampling variance for the log of FHB index
fhb_sev$vi <- with(fhb_sev, V_sev / (n * sev^2))

# Sampling variance for the log of yield
fhb_yield$vi <- with(fhb_yield, V_yield / (n * yield^2))

# Sampling variance for yield
fhb_yield$vi2 <- fhb_yield$V_yield/fhb_yield$n # multivariate approach
```

## Treatments

There are tree fungicides of interest in this study:

- Tebuconazole
- Propiconazole
- Carbendazim

Most of them were used as a single application but Tebuconazole was tested as single or sequential application. Let's get the number of entries per fungicide. The `AACHECK` refers to the plots where no fungicide was applied, or the control.

```{r}
table(fhb_sev$AI, fhb_sev$n_spray2)
```

Let's create another variable to identify each of these four treatments. We will omit CARB applied once because there are only three entries.

```{r}
library(tidyverse)
fhb_sev <- fhb_sev %>% 
  mutate(AI_nspray2 = paste(AI, n_spray2,sep='_')) %>% 
  filter(AI_nspray2 != "CARB_1") %>% 
  filter(trial, length(trial)>1)

fhb_yield <- fhb_yield %>% 
  mutate(AI_nspray2 = paste(AI, n_spray2,sep='_')) %>% 
  filter(AI_nspray2 != "CARB_1") %>% 
  filter(trial, length(trial)>1)

```

Finally, we create a unique identifier for the treatment within a trial. This can be used further as random effects in the model.

```{r}
# ID variable for each entry within a trial

fhb_sev$id <- 1:nrow(fhb_sev)
fhb_yield$id <- 1:nrow(fhb_yield)
```

## Descriptive analysis

We will get some summary information which may of use to describe the dataset when writing the material and methods and results section of the paper.

```{r}
# Number of entries by fungicide and number of sprays
table(fhb_yield$AI, fhb_yield$n_spray2)

# Number of unique trials
length(unique(fhb_yield$trial))

# Number of unique studies
length(unique(fhb_yield$study))

# Number of trials by year
fhb_trial <- fhb_yield %>% 
  group_by(trial) %>% 
  filter(row_number() ==1)

data.frame(table(fhb_trial$year))

nrow(table(fhb_trial$year))

# Number of trial by location
data.frame(table(fhb_trial$location))

# Number of trials by publication type
data.frame(table(fhb_trial$publication))

# Number of trials by state
data.frame(table(fhb_trial$state))

# n. of entries with D value lower than zero
negat_D <- fhb_yield %>% 
  filter(D < 0)
```

We also want to check whether the difference in means (D) can be expressed in absolute rather than relative terms. The former is indicated for situations in which the coefficient of variation for the response variable in the check treatment is not too high (Paul et al. 2008)

```{r eval=FALSE, include=FALSE}
# CV for FHB index
## All entries
cv_index_all <- (sd(fhb_sev$sev)/mean(fhb_sev$sev))*100

## For control only
cv_index_check <- (sd(fhb_sev$sev_check)/mean(fhb_sev$sev_check))*100

### CV for yield

## All entries
cv_yld_all <- (sd(fhb_yield$yield)/mean(fhb_yield$yield))*100

## For control only
cv_yld_check <- (sd(fhb_yield$yield_check)/mean(fhb_yield$yield_check))*100

```

We can see that CV was high for FHB index, but much lower for yield. It is safer then to work with the absolute means difference for yield than FHB index, which will also require log-transformation.

## Meta-analysis

### Control efficacy

Let's fit an arm-based model using the `rma.mv` function of metafor to get the estimates for each treatment.

```{r, message=FALSE, warning=FALSE}
fhb_mv_AI <- rma.mv(yi, vi, 
                    mods = ~ AI_nspray2, 
                    method="ML",
                    random = list(~ AI_nspray2 | trial, ~1 | id), 
                    struct="UN", data=fhb_sev)

```

Let's see the summary ouput provided by this function. Note that the intercept is the control treatment and all treatments differed significantly from the control based on the very low P-values.

```{r}
fhb_mv_AI
```

We can set linear contrasts between treatments of interest and get the P-valued using the `anova` function.

```{r}
anova(fhb_mv_AI, L=rbind(c(0,1,-1,0,0), 
                         c(0,0,-1,1,0),
                         c(0,0,-1,0,1),
                         c(0,0,0,-1,1)))

```

Let's back-transform (exponential) the estimated mean (and respective 95%CIs) difference of the logs to obtain the response ratio, which is further expressed as percent reduction in FHB index relative to the check treatment (control efficacy) for each fungicide. We then prepare the data for plotting purposes.

```{r}
results_AI <- data.frame(cbind((1-exp(fhb_mv_AI$b))*100, 
                              (1-exp(fhb_mv_AI$ci.lb))*100,
                              (1-exp(fhb_mv_AI$ci.ub))*100))

treat <- c("Check", "CARB2x", "PROP2x", "TEBU1x", "TEBU2x")
efficacy <- tbl_df(results_AI)
efficacy$treat <- treat
efficacy$se <- fhb_mv_AI$se
colnames(efficacy) <- c("Mean", "CIsup", "CIinf", "Treatment", "SE")
efficacy
```

We can visualize this quickly, but we omit the check treatment.

```{r}
efficacy %>% 
  filter(Treatment != "Check") %>% 
  ggplot(aes(Treatment, Mean))+
  geom_point(aes(size=1/SE), shape=15)+
  geom_linerange(aes(ymin = CIinf, ymax = CIsup))+
  coord_flip()+
  theme_grey()+
  ylim(20,80)
```

#### Inconsistency: index

In a network meta-analysis, it is important to check whether the results are consistent across the different designs. We included this variable manually in the original dataset and tested the interaction effect of treatments and design group.

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}

### Number of trials per design (cons_group)
table(fhb_sev$cons_group)

### Model testing interaction treat x design
fhb_mv_AI_cons <- rma.mv(yi, vi, 
                         mods = ~ AI_nspray2 * cons_group, 
                         method="ML",
                         random = ~ AI_nspray2 | trial/cons_group, 
                         struct="UN", 
                         data=fhb_sev)
anova(fhb_mv_AI_cons, btt=7:10)

```

#### Moderators

##### FHB index base

```{r, message=FALSE, warning=FALSE}
# create the binary variable (dis_press)
fhb_sev$dis_press <- ifelse(as.numeric(fhb_sev$sev_check) >7.0, 2, 1)

# summarize number of trials per category of dis_press
table(fhb_sev$AI_nspray2, fhb_sev$dis_press)

# Test effect of moderator 
fhb_mv_AI_sev_check <- rma.mv(yi, vi, 
                              mods = ~ AI_nspray2*factor(dis_press),
                              method="ML",
                              random = list(~ AI_nspray2 | trial, ~1 | id),
                              struct="UN", 
                              data=fhb_sev)

# Contrast levels of moderators
anova(fhb_mv_AI_sev_check,btt=7:10)

```

##### Yield base

```{r, message=FALSE, warning=FALSE}

# create binary variable (yield_class)
summary(fhb_sev$yield_check) # Median = 2993; Mean = 2915
fhb_sev$yield_class <- ifelse(fhb_sev$yield_check > 3000, 2, 1)

# check number of trials per yield_class category
table(fhb_sev$yield_class, fhb_sev$AI_nspray2)

# test the effect of moderator
fhb_mv_AI_yield_class <- rma.mv(yi, vi, 
                                mods = ~ AI_nspray2*factor(yield_class),
                                method="ML",
                                random = list(~ AI_nspray2 | trial, ~1 | id), 
                                struct="UN", 
                                data=fhb_sev)

anova(fhb_mv_AI_yield_class, btt=7:10)

```

##### Year as continuous

```{r, message=FALSE, warning=FALSE}
# Moderator year (continuous)
# Number of entries by fungicide and year
table(fhb_sev$AI_nspray2, fhb_sev$year)

fhb_mv_AI_year <- rma.mv(yi, vi, 
                         mods = ~ AI_nspray2*as.numeric(year), 
                         method="ML",
                         random =list(~ AI_nspray2 | trial, ~1 | id),
                         struct="UN", 
                         data=fhb_sev)

anova(fhb_mv_AI_year, btt=7:10)

```

### Yield response: relative

```{r}
### AI_nspray2 (AI and number of sprays)

fhb_mv_AI <- rma.mv(yi, vi, 
                    mods = ~ AI_nspray2, 
                    method="ML",random = list(~ AI_nspray2 | trial, ~1 | id), 
                    struct="UN", data=fhb_yield)

summary(fhb_mv_AI)

results_I<- data.frame(cbind((exp(fhb_mv_AI$b)-1)*100, 
                             (exp(fhb_mv_AI$ci.lb)-1)*100,
                             (exp(fhb_mv_AI$ci.ub)-1)*100))

results_I

# Linear contrasts between treatments

anova(fhb_mv_AI, L=rbind(c(0,1,-1,0,0), 
                         c(0,0,1,-1,0),
                         c(0,0,1,0,-1),
                         c(0,0,0,-1,1)))

```

### Yield response: absolute

```{r}
### by fungicide treatment

fhb_mv_AI_D <- rma.mv(yield, vi2,
                      mods = ~ AI_nspray2, 
                      method="ML",random = list(~ AI_nspray2 | trial, ~1 | id),                       struct="UN", 
                      data=fhb_yield, 
                      control = list(optimizer="nlm"))
summary(fhb_mv_AI_D)

# Linear contrasts between treatments

anova(fhb_mv_AI_D, L=rbind(c(0,1,-1,0,0), 
                         c(0,0,1,-1,0),
                         c(0,0,1,0,-1),
                         c(0,0,0,-1,1)))
```

#### Inconsistency

```{r}

### Groups were inclued directly in the spreadsheet
table(fhb_yield$cons_group)

### Model 
fhb_mv_AI_cons <- rma.mv(yield, vi2, 
                         mods = ~ AI_nspray2 * cons_group, 
                         method="ML",
                         random = ~ AI_nspray2 | trial/cons_group, 
                         struct="UN", data=fhb_yield)

summary(fhb_mv_AI_cons)

results_I<- data.frame(cbind((exp(fhb_mv_AI$b)-1)*100, 
                             (exp(fhb_mv_AI$ci.lb)-1)*100,
                             (exp(fhb_mv_AI$ci.ub)-1)*100))
results_I

#anova(fhb_mv_AI_cons, btt=14:26)

```