### 2017-07-24 index.Rmd

---
title: ''
output:
  html_document:
    css: my-style.css
    theme: flatly
    toc: true
    toc_float: 
     collapsed: false
     smooth_scroll: false
---

## About

<img align=right width=300 src = "figures/article_cover.png">

This webpage aims at providing access to the full text, the raw data and the `R` codes for reproducing the analysis and results in an article from a masters research conducted in the  [Epidemiology Lab](http://www.dfp.ufv.br/epidemiology) which was published in Plant Disease [(Machado et al. 2017)](http://apsjournals.apsnet.org/doi/abs/10.1094/PDIS-03-17-0340-RE). 
We used meta-analysis, preceeded by a systematic review to identify and select primary studies, to summarize the effects of specific fungicide  (two triazoles and one benzimidazole) for controling Fusarium head blight, a fungal disease that damages the grain and produces mycotoxins that are harmfull to consumers.

It is intended to demonstrate all the decisions taken over the course of the analysis the data, from data import to transform, calculation of new variables, modeling and preparation of high-quality plots for publication. It is hoped that network meta-analysis, whenever multiple-treatments are under comparison, is better understood and used more extensively whenever appropriate. Navigate using the top menu to get access to the main four sections.